MATH 3610 - Project \#1

### 2015-09-26 proj1.tex
adding disease spread with vaccine and with no vaccine

\documentclass{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{amsmath}

@@

\section{Evaluation Models}

@@

>>>>

\documentclass{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{tikz}

@@

\section{Evaluating Models}

@@

\section{Simple Continuous Model}

First, we propose a deterministic model which doesn't divide the population 
	into sections at all.
This model will help give us an idea of what the effect of dividing
	the population into sections will be.
In addition, it can give us an idea of how vaccination affects the spread of
	disease within a population.

\subsection{Disease Spread with No Vaccine}

We include this derivation of the logistic model for disease spread
	to compare with our derivation of the modified logistic model
	for disease spread and vaccination.

Suppose there are two groups of people, with $S$ representing the number of 
	sick people, and $H$ represnting the number of healthy people.
Suppose the probability of a single healthy person becoming sick is proportional
	only to the total number of sick people.
Since there are $H$ healthy people, the expected number of healthy people
	who become sick should be proportional to $S H$.
This model does not take into account recovery from sickness.
If we introduce a parameter $r$, we can express this relationship as a system
	of differential equations:

\begin{align*}
\frac{dH}{dt} & = - r S H \\
\frac{dS}{dt} & = r S H \\
\end{align*}

Notice that the total number of people, $S + H$, is conserved, i.e.,

\[ \frac{d}{dt} \left( S + H \right) = \frac{dS}{dt} + \frac{dH}{dt}
	= r S H - r S H = 0 \]

If we introduce another parameter, $K$, to represent the total population,
	such that $S + H = K$ for all time, then the differential equations
	can be decoupled:

\begin{align*}
\frac{dH}{dt} & = - r (K - H) H \\
\frac{dS}{dt} & = r S (K - S) \\
\end{align*}

These are two independant logistic models.

\subsection{Diease Spread with Vaccine}

Suppose we introduce a third group, $V$, representing people who are vaccinated	
	against the disease.
Suppose we randomly vaccinate members of the healthy population,
	a constant number each time step.
This introduces a third parameter, $r_v$, the rate of vaccinations.
The vaccinated population isn't seceptible to the disease.
This process is modeled by the following system of diffential equations:

\begin{align*}
\frac{dS}{dt} & = r S H \\
\frac{dH}{dt} & = - r S H - r_v \\
\frac{dV}{dt} & = r_v
\end{align*}

This model is accurate only if all three of the populations remain
	positive.
This is important to remember.
The number of healthy people will always be decreasing, so once
	the number of healthy people hits zero, we stop the model.
The sick people stay sick, and the vaccinated people stay vaccinated.

The number of vaccinated people as a function of time is:

\[ V(t) = r_v t \]

Again, we find the conserved quantity $\left( S + H + V \right)$ and
	name it $K$.
For the model to be valid, we need $H = K - S - V \geq 0$.
Therefore, the model is valid as long as:

\[ K \geq S + r_v t \]

We have now made these differential equations independant:

\begin{align*}
\frac{dS}{dt} & = r S \left( K - S - r_v t \right)\\
\frac{dH}{dt} & = - r H \left( K - H - r_v t \right) - r_v \\
\frac{dV}{dt} & = r_v t
\end{align*}
